% Thesis Acknowledgements ------------------------------------------------


%\begin{acknowledgementslong} %uncommenting this line, gives a different acknowledgements heading
\begin{acknowledgements}      %this creates the heading for the acknowlegments
Many thanks are due to the people that got me to where I am. For as long as I can remember, I have aspired to be an astrophysicist and while I have never lost that passion and fire for astonomy, it would not remain so resolute without the unwavering support of those around me. To anyone I have missed in this list, I sincerely apologise, but know that I thank you sincerely for helping me become Dr O'Ryan: astrophysicist.
\\
\\
% For Brooke
First, I offer my sincere gratitude to my supervisor Brooke Simmons. Your encouragement, guidance, enthusiasm and support over these many years have made this possible.
\\
\\
% For collaborators
Bruno Merin, Chris Lintott, Mike Walmsley, Becky Smethurst, Tobias Geron, Sandor Kruk, Kameswara Mantha, Karen Masters, and the whole Galaxy Zoo Team who were all so supportive, insightful and inspiring throughout my PhD. 
\\
\\
% Those at ESA
The ESA trainee cohort of 2022. When my love for astronomy was wavering and committment to academia in doubt you showed me a wonderful, inquisitive and welcoming environment that brought that passion back. I want to also thank Karolin Frohnapfel, Emma Vellard, Aurelian Verdier, Antonia Vojtekova, Anna Anku and Marie Scholch specifically from this cohort. I went alone to a new, big city and, with your help, left it with some of the best memories of my life.  
\\
\\
% Those at 1715Labs
The team at 1715Labs. Working with you was a brilliant experience, and gave me a peek behind the curtain into life in industry.
\\
\\
% Those in the office
Those in the Lancaster PhD office, who made it such a welcoming place. To Jamie Dumayne and Izzy Garland for many fun pub trips, and to Heather Wade and Tom Cornish for hosting some awesome parties. To Rahul Rana, Pascale Desmet, Jon Carrick, Nick Amos, Harry Stevenson and Andrew Milligan for some great laughs, scientific discussion and knowledge about the world.
\\
\\
% Matt, Amy and Nikita.
To my closest friends in Lancaster: Amy Hewitt, Matthew Thorne, Nikita Mehta and Zach Mason. You got me through Covid, and some of the toughest times of my PhD as we all went crazy together in those houses. While you're all scattered to the wind, doing and excelling at your own thing, I hope that we will meet again.
\\
\\
% For my parents
Most importantly, to my parents. You fanned those initial flames that gave me a passion for astronomy that has never left me, and driven me since I was a child. Your support then, and now, has been a rock for me and it is only because of you that I am where I am today.
\\
\\
% For my brother and sister
To my brother and sister, Michael and Ruth. For putting up with my incessent chat about astronomy and for always being there.
\\
\\
% To Calum & Ronan
To Calum Cooper and Ronan Duff, who have supported me completely throughout my life. Even when my decisions lead me to leave you in Glasgow, you stood by me and still remain my closest friends to this day.
\\
\\
% Lancaster University
This research was funded by the Science and Technology Council of the UK, without them this work would not have been possible. I also thank my host intitustion: Lancaster University.
\\
\\
Finally, to Abby. While we step into the unknown of what happens next, know you were my rock through the hardest points of my PhD. Your unwavering support and selflessness has left an impact on me that will persist for the rest of my life.




\end{acknowledgements}
%\end{acknowledgmentslong}

% ------------------------------------------------------------------------


